Distribution of dissimilarity and breadth indices within each dataset

### 2021-03-22 analysis/rev_prototyping/draft_reports/appendices/s6_breadth_index_fig.Rmd
rev draft

---
title: "Figures and results for main manuscript"
author: Renata Diaz
date: "`r Sys.Date()`"
output: 
    github_document:
       df_print: kable
       toc: true
---

@@

# Datasets by S and N (Figure 1)

```{r, fig.dim = c(7,7)}

s4_fig1 <- ggplot(all_di, aes(x = s0, y = n0, color = Dataset)) +
  geom_point(alpha = .1) +
  geom_point(data = filter(all_di), alpha = .4) +
  theme_bw() +
  scale_color_viridis_d(end = .9) +
  ggtitle("Communities by S and N") +
  xlab("Species richness (S); note log scale") +
  ylab("Total abundance (N); note log scale") +
  scale_x_log10() +
  scale_y_log10()

plot(s4_fig1)
```

# Illustrations of 95% interval (Figure 2)

To show the 95% interval, we need to load the distribution of shape metric values from the samples from the feasible set for a few communities. See rov_metric.md.

<!-- ```{r, fig.dim = c(7,7)} -->
<!-- library(drake) -->

<!-- db <- DBI::dbConnect(RSQLite::SQLite(), here::here("analysis", "drake", "drake-cache-net.sqlite")) -->
<!-- cache <- storr::storr_dbi("datatable", "keystable", db) -->
<!-- cache$del(key = "lock", namespace = "session") -->

<!-- net_summary <- readd(all_di_summary, cache = cache) -->

<!-- net_summary <- net_summary %>% -->
<!--   mutate(log_nparts = log(gmp:::as.double.bigz(nparts))) -->

<!-- example_fs <- readd(fs_s_44_n_13360, cache = cache) -->

<!-- example_di <- readd(di_fs_s_44_n_13360, cache =cache) -->

<!-- example_fs <- example_fs %>% -->
<!--   left_join(example_di) %>% -->
<!--   left_join(net_summary) -->

<!-- example_fs2 <- readd(fs_s_13_n_315, cache = cache) -->

<!-- example_di2 <- readd(di_fs_s_13_n_315, cache =cache) -->

<!-- example_fs2 <- example_fs2 %>% -->
<!--   left_join(example_di2) %>% -->
<!--   left_join(net_summary) -->

<!-- example_fs3 <- readd(fs_s_4_n_34, cache = cache) -->

<!-- example_di3 <- readd(di_fs_s_4_n_34, cache =cache) -->

<!-- example_fs3 <- example_fs3 %>% -->
<!--   left_join(example_di3) %>% -->
<!--   left_join(net_summary) -->

<!-- breadth_plots <- list( -->

<!--   ggplot(example_fs3, aes(rank, abund, group = sim, color = skew)) + -->
<!--     geom_line(alpha = .25) + -->
<!--     theme_bw() + -->
<!--     scale_color_viridis_c(option = "plasma", end = .8) + -->
<!--     ggtitle("Small community", subtitle = paste0("S = ", (example_fs3$s0), "; N = ", (example_fs3$n0[1]))) + -->
<!--     theme(legend.position = "right") + -->
<!--     xlab("Rank") + -->
<!--     ylab("Abundance"), -->

<!--   ggplot(example_fs3, aes(skew)) + -->
<!--   #  geom_density() + -->
<!--     geom_histogram(bins = 50) + -->
<!--     theme_bw() + -->
<!--     geom_vline(xintercept = c(example_fs3$skew_97p5[1], example_fs3$skew_2p5[1]), color = "red") + -->
<!--     ggtitle("", subtitle = paste0("Breadth index: ", round((example_fs3$skew_95_ratio_2t[1]), 2))) + -->
<!--     xlab("Skewness") + -->
<!--     ylab("Count"), -->

<!--   ggplot(example_fs2, aes(rank, abund, group = sim, color = skew)) + -->
<!--     geom_line(alpha = .1) + -->
<!--     theme_bw() + -->
<!--     scale_color_viridis_c(option = "plasma", end = .8) + -->
<!--     ggtitle("Medium community", subtitle = paste0("S = ", (example_fs2$s0), "; N = ", (example_fs2$n0[1]))) + -->
<!--     theme(legend.position = "right")+ -->
<!--     xlab("Rank") + -->
<!--     ylab("Abundance") + -->
<!--     ylim(0, 200), # Remove 3 sads that make the axes too big to be interpretable -->

<!--   ggplot(example_fs2, aes(skew)) + -->
<!--   #  geom_density() + -->
<!--     geom_histogram(bins = 50) + -->
<!--     theme_bw() + -->
<!--     geom_vline(xintercept = c(example_fs2$skew_97p5[1], example_fs2$skew_2p5[1]), color = "red") + -->
<!--     ggtitle("", subtitle =  paste0("Breadth index: ", round((example_fs2$skew_95_ratio_1t[1]), 2)))+ -->
<!--     xlab("Skewness") + -->
<!--     ylab("Count"), -->

<!--   ggplot(example_fs, aes(rank, abund, group = sim, color = skew)) + -->
<!--     geom_line(alpha = .1) + -->
<!--     theme_bw() + -->
<!--     scale_color_viridis_c(option = "plasma", end = .8) + -->
<!--     ggtitle("Large community", subtitle = paste0("S = ", (example_fs$s0), "; N = ", (example_fs$n0[1]))) + -->
<!--     theme(legend.position = "right") + -->
<!--     ylim(0, 4000) + # Remove a very few very very uneven SADs that make the scale too big to be interpretable -->
<!--     theme(axis.text.y = element_text(size = 6, angle = 60))+ -->
<!--     xlab("Rank") + -->
<!--     ylab("Abundance"), -->

<!--   ggplot(example_fs, aes(skew)) + -->
<!--    # geom_density() + -->
<!--     geom_histogram(bins = 50) + -->
<!--     theme_bw() + -->
<!--     geom_vline(xintercept = c(example_fs$skew_97p5[1], example_fs$skew_2p5[1]), color = "red") + -->
<!--     ggtitle("", subtitle =  paste0("Breadth index: ", round((example_fs$skew_95_ratio_1t[1]), 2)))+ -->
<!--     xlab("Skewness") + -->
<!--     ylab("Count") -->

<!-- ) -->

<!-- fig_1 <- gridExtra::grid.arrange(grobs = breadth_plots, ncol = 2, top = textGrob("Figure 1", gp = gpar(fill = "white"))) -->

<!-- plot(fig_1) -->

<!-- DBI::dbDisconnect(db) -->
<!-- rm(cache) -->
<!-- rm(db) -->
<!-- ``` -->

# Dissimilarity (Supplement)

```{r, fig.dim = c(3, 6)}
plot_percentile_hist <- function(di_df, col_name, plot_name, tails = 2, facetvar = "Dataset", min_s0 = 0) {

  if(tails == 2) {
    cutoff_percentiles = c(2.5, 97.5)
    min_nparts = 40
  } else if (tails== 1) {
    cutoff_percentiles = c(95)
    min_nparts = 20
  }

  ggplot(filter(di_df, nparts > min_nparts, s0 >= min_s0), aes(response)) +
    geom_histogram(bins = 40, boundary = 100) +
    theme_bw() +
    xlab("") +
    ylab("") +
    geom_vline(xintercept = cutoff_percentiles, color = "red") +
    ggtitle( plot_name) +
    facet_wrap(vars(facetvar), ncol = 1, scales = "free_y")+ 
    theme(plot.title = element_text(size=10))

all_di <- all_di %>%
  mutate(`Observed \npercentile score` = ifelse(real_po_percentile_excl > 95, "High (> 95)", "Less than 95"))

```{r, fig.dim = c(4, 4)}
ggplot(all_di, aes(sim_pos_from_best, real_po, color = `Observed \npercentile score`)) +
  geom_point(alpha = .3) + 
  geom_point(data = filter(all_di, s0 < 0)) +
  geom_abline(slope = 1, intercept = 0) +
  scale_color_viridis_d(end = .8, direction = -1) +
  xlab("Mean of elements from feasible set") +
  ylab("Observed value") +
  ggtitle("Dissimilarity to the central tendency") +  theme(legend.position = "bottom")#,
# 
# 
# ggplot(filter(all_di, nparts < 10 ^ 50), aes(nparts, sim_pos_from_best, color = Dataset)) +
#   geom_point(alpha = .3) + 
#   geom_point(data = filter(all_di, s0 < 0)) +
#   scale_color_viridis_d(end = .9) +
#   xlab("Number of elements in the feasible set") +
#   ylab("Mean dissimilarity of feasible set to central tendency") +
#   scale_x_log10() +
#   ggtitle("Over the size of the feasible set")
# ), ncol = 1,
#top = textGrob("Dissimilarity to central tendency"))

# Metric histograms by dataset (Figure 4)

```{r first hists, fig.dim = c(10,5)}

  plot_percentile_hist(all_di, "real_po_percentile_mean", "Dissimilarity to the \ncentral tendency", facetvar = "Dataset", tails = 1),
  plot_percentile_hist(all_di, "nsingletons_percentile_mean", "\nNumber of rare species", facetvar = "Dataset"),
  plot_percentile_hist(all_di, "skew_percentile_mean", "\nSkewness", facetvar = "Dataset", min_s0 = 3),
  plot_percentile_hist(all_di, "simpson_percentile_mean", "\nSimpson evenness", facetvar = "Dataset"),
  plot_percentile_hist(all_di, "shannon_percentile_mean", "\nShannon diversity", facetvar = "Dataset")
), ncol = 5,
top = textGrob("Figure 2", gp = gpar(fill = "white")), left = textGrob("Number of communities", rot = 90, gp = gpar(fill = "black")),
bottom = textGrob("Percentile rank of observed value relative to feasible set"), gp = gpar(fill = "white"))
```

# Results table

```{r}
diss_results <- all_di %>%
  filter(nparts > 20) %>%
  group_by(Dataset) %>%
  summarize(high_diss = mean(real_po_percentile_excl > 95),
            nsites = dplyr::n()) %>%
  ungroup() %>%
  mutate(`High dissimilarity` = paste0(signif(100 * high_diss, 2), "%; n = ", nsites)) %>%
  select(-nsites)

skew_results <- all_di %>%
  filter(nparts > 40, s0 > 2) %>%
  group_by(Dataset) %>%
  summarize(high_skew = mean(skew_percentile_excl > 97.5),
            low_skew = mean(skew_percentile < 2.5),
            nsites = dplyr::n()) %>%
  ungroup() %>%
  mutate(`High skew` = paste0(signif(100 * high_skew, 2), "%; n = ", nsites),
         `Low skew` = paste0(signif(100 * low_skew, 2), "%; n = ", nsites)) %>%
  select(-nsites)

metric_results <- all_di %>%
  filter(nparts > 40) %>%
  group_by(Dataset) %>%
  summarize(high_simpson = mean(simpson_percentile_excl > 97.5),
            low_simpson = mean(simpson_percentile < 2.5),
            high_shannon = mean(shannon_percentile_excl > 97.5),
            low_shannon = mean(shannon_percentile < 2.5),
            high_nsingletons = mean(nsingletons_percentile_excl > 97.5),
            low_nsingletons = mean(nsingletons_percentile < 2.5),
            nsites = dplyr::n()) %>%
  ungroup() %>%
  mutate(`High Simpson` = paste0(signif(100 * high_simpson, 2), "%; n = ", nsites),
         `Low Simpson` = paste0(signif(100 * low_simpson, 2), "%; n = ", nsites),
         `High Shannon` = paste0(signif(100 * high_shannon, 2), "%; n = ", nsites),
         `Low Shannon` = paste0(signif(100 * low_shannon, 2), "%; n = ", nsites),
         `High proportion of rare species` = paste0(signif(100 * high_nsingletons, 2), "%; n = ", nsites),
         `Low proportion of rare species` = paste0(signif(100 * low_nsingletons, 2), "%; n = ", nsites)) %>%
  select(-nsites)

all_results <- left_join(diss_results, skew_results) %>%
  left_join(metric_results)

cols1 <- c("Dataset", "High dissimilarity", "High proportion of rare species", "High skew", "Low Simpson", "Low Shannon")

all_results %>%
  select(cols1)

cols2 <- c("Dataset", "Low proportion of rare species", "Low skew", "High Simpson", "High Shannon")

all_results %>%
  select(cols2)

# Narrowness

@@

  ggplot(di_df, aes(nparts, response, color = Dataset)) +
    geom_point(alpha = .3) +
    geom_point(data = filter(di_df, s0 < 0)) +
    scale_color_viridis_d(end = .9) +
    xlab("") +
    ylab(ylabel) +
    scale_x_log10() +
    ggtitle(yvar_name) + 
    theme(legend.position = "none")+ 
    theme(plot.title = element_text(size=10))
}

@@

```{r, fig.dim = c(7,4)}

gridExtra::grid.arrange(grobs = list(
  plot_narrowness(all_di, "sim_pos_from_best", "Dissimilarity \nto central tendency") + ylab("Mean dissimilarity") + theme(legend.position = "bottom", legend.direction = "vertical" , legend.margin = margin(b = .25, r = 0, t = 0, l = 0, unit = "in")), 
  gridExtra::grid.arrange(grobs = list(
    plot_narrowness(all_di, "skew_95_ratio_2t", "\nSkewness"),
    plot_narrowness(all_di, "simpson_95_ratio_2t", "\nSimpson evenness")+ ylab(""),
    plot_narrowness(all_di, "nsingletons_95_ratio_2t", "\nNumber of rare species"),
       plot_narrowness(all_di, "shannon_95_ratio_2t", "\nShannon diversity") + ylab("")
  ), ncol = 2)), ncol = 2, widths = c( 2, 4), bottom = textGrob("Number of elements in the feasible set"))

@@

# Gentry

gentry_di <- filter(all_di, Dataset == "Gentry")

gentry_di <- gentry_di %>%
  mutate(low_avg_abund = ifelse(n0/s0 < 3, "N:S < 3", "N:S > 3"))

  plot_percentile_hist(gentry_di, "nsingletons_percentile_mean", "\nNumber of rare species", facetvar = "low_avg_abund"),
  plot_percentile_hist(gentry_di, "skew_percentile_mean", "\nSkewness", facetvar = "low_avg_abund", min_s0 = 3),
  plot_percentile_hist(gentry_di, "simpson_percentile_mean", "\nSimpson evenness", facetvar = "low_avg_abund"),
  plot_percentile_hist(gentry_di, "shannon_percentile_mean", "\nShannon diversity", facetvar = "low_avg_abund")
), ncol = 2,
 left = textGrob("Number of communities", rot = 90, gp = gpar(fill = "black")),
bottom = textGrob("Percentile rank of observed value relative to feasible set"), gp = gpar(fill = "white"))

# Comparison of FIA and comparably sized communities (Figure 4)

```{r, fig.dim = c(10, 5)}

fia_max_s = max(filter(all_di, dat == "fia")$s0)
fia_max_n = max(filter(all_di, dat == "fia")$n0)

fia = filter(all_di, dat == "fia")

small_di <- all_di %>%
  filter(s0 <= fia_max_s,
         n0 <= fia_max_n) %>%
  mutate(fia_yn = ifelse(dat == "fia", "fia", "other datasets"),
         right_corner = FALSE)

for(i in 1:nrow(small_di)) {
  if(small_di$dat[i] != "fia") {
    fia_match_s0 = filter(fia, s0 >= small_di$s0[i])

    max_n0_for_this_s0 = max(fia_match_s0$n0)

    small_di$right_corner[i] = small_di$n0[i] > max_n0_for_this_s0

  }
}

small_di <- small_di %>%
  filter(!right_corner)
not_fia <- filter(small_di, fia_yn != "fia")
fia = filter(small_di, dat == "fia")

small_di_s_n <- small_di %>%
  select(s0, n0) %>%
  distinct() %>%
  mutate(in_fia = FALSE,
         in_not_fia = FALSE)

for(i in 1:nrow(small_di_s_n)) {

  this_s0 = small_di_s_n$s0[i]
  this_n0 = small_di_s_n$n0[i]

  fia_matches = filter(fia, s0 == this_s0, n0 == this_n0)
  not_fia_matches = filter(not_fia, s0 == this_s0, n0 == this_n0)

  small_di_s_n$in_fia[i] = nrow(fia_matches) > 0
  small_di_s_n$in_not_fia[i] = nrow(not_fia_matches) > 0

small_di_s_n <- filter(small_di_s_n, in_fia, in_not_fia)

set.seed(1977)

subsamples <- list()

for(i in 1:nrow(small_di_s_n)) {

  this_s0 = small_di_s_n$s0[i]
  this_n0 = small_di_s_n$n0[i]

  fia_matches = filter(fia, s0 == this_s0, n0 == this_n0)

  not_fia_matches = filter(not_fia, s0 == this_s0, n0 == this_n0)

  n_to_draw = min(nrow(fia_matches), nrow(not_fia_matches))

  fia_draw = sample.int(n = nrow(fia_matches), size = n_to_draw, replace = F)
  not_fia_draw = sample.int(n = nrow(not_fia_matches), size = n_to_draw, replace = F)

  subsamples[[i]] <- (bind_rows(fia_matches[fia_draw,], not_fia_matches[not_fia_draw, ]))

sub_di <- bind_rows(subsamples) %>%
  mutate(Dataset = ifelse(fia_yn == "fia", "FIA", "Other datasets"))

fia_di <- filter(sub_di, fia_yn == "fia")
other_di <- filter(sub_di, fia_yn != "fia")

ks_compare <- function(fia_df, other_df, compare_var) {

  if(grepl("skew", compare_var)) {
    fia_df <- filter(fia_df, s0 > 2)
    other_df <- filter(fia_df, s0 >2)
  }

  if(grepl("sim_pos", compare_var)) {
    fia_df <- filter(fia_df, nparts > 20)
    other_df <- filter(other_df, nparts > 20)
  } else {
      fia_df <- filter(fia_df, nparts > 40)
    other_df <- filter(other_df, nparts > 40)
  }

  kstest <- ks.test(fia_df[[compare_var]], other_df[[compare_var]])

  return(data.frame(
    var = compare_var,
    d = kstest$statistic,
    p = kstest$p.value
  ))
}

simpson_rat_ks <- ks_compare(fia_di, other_di, "simpson_95_ratio_2t")
skew_rat_ks <- ks_compare(fia_di, other_di, "skew_95_ratio_2t")
shannon_rat_ks <- ks_compare(fia_di, other_di, "shannon_95_ratio_2t")
nsingletons_rat_ks <- ks_compare(fia_di, other_di, "nsingletons_95_ratio_2t")
diss_rat_ks <- ks_compare(fia_di, other_di, "sim_pos_from_best")

bind_rows(simpson_rat_ks, skew_rat_ks, shannon_rat_ks, nsingletons_rat_ks, diss_rat_ks)

simpson_perc_ks <- ks_compare(fia_di, other_di, "simpson_percentile")
skew_perc_ks <- ks_compare(fia_di, other_di, "skew_percentile_excl")
shannon_perc_ks <- ks_compare(fia_di, other_di, "shannon_percentile")
nsingletons_perc_ks <- ks_compare(fia_di, other_di, "nsingletons_percentile_excl")
diss_perc_ks <- ks_compare(fia_di, other_di, "real_po_percentile_excl")

bind_rows(simpson_perc_ks, skew_perc_ks, shannon_perc_ks, nsingletons_perc_ks, diss_perc_ks)

  plot_percentile_hist(sub_di, "real_po_percentile_mean", "Dissimilarity to the \ncentral tendency", facetvar = "Dataset", tails = 1),
  plot_percentile_hist(sub_di, "nsingletons_percentile_mean", "\nNumber of rare species", facetvar = "Dataset"),
  plot_percentile_hist(sub_di, "skew_percentile_mean", "\nSkewness", facetvar = "Dataset", min_s0 = 3),
  plot_percentile_hist(sub_di, "simpson_percentile_mean", "\nSimpson evenness", facetvar = "Dataset"),
  plot_percentile_hist(sub_di, "shannon_percentile_mean", "\nShannon diversity", facetvar = "Dataset")
), ncol = 5,
top = textGrob("Figure 2", gp = gpar(fill = "white")), left = textGrob("Number of communities", rot = 90, gp = gpar(fill = "black")),
bottom = textGrob("Percentile rank of observed value relative to feasible set"), gp = gpar(fill = "white"))

  plot_narrowness_hist(sub_di, "sim_pos_from_best", "Dissimilarity to the \ncentral tendency", facetvar = "Dataset"),
  plot_narrowness_hist(sub_di, "nsingletons_95_ratio_2t", "\nNumber of rare species", facetvar = "Dataset"),
  plot_narrowness_hist(sub_di, "skew_95_ratio_2t", "\nSkewness", facetvar = "Dataset", min_s0 = 3),
  plot_narrowness_hist(sub_di, "simpson_95_ratio_2t", "\nSimpson evenness", facetvar = "Dataset"),
  plot_narrowness_hist(sub_di, "shannon_95_ratio_2t", "\nShannon diversity", facetvar = "Dataset")
), ncol = 5,
 left = textGrob("Number of communities", rot = 90, gp = gpar(fill = "black")), gp = gpar(fill = "white"))

diss_results <- sub_di %>%
  filter(nparts > 20) %>%
  group_by(Dataset) %>%
  summarize(high_diss = mean(real_po_percentile_excl > 95),
            nsites = dplyr::n()) %>%
  ungroup() %>%
  mutate(`High dissimilarity` = paste0(signif(100 * high_diss, 2), "%; n = ", nsites)) %>%
  select(-nsites)

skew_results <- sub_di %>%
  filter(nparts > 40, s0 > 2) %>%
  group_by(Dataset) %>%
  summarize(high_skew = mean(skew_percentile_excl > 97.5),
            low_skew = mean(skew_percentile < 2.5),
            nsites = dplyr::n()) %>%
  ungroup() %>%
  mutate(`High skew` = paste0(signif(100 * high_skew, 2), "%; n = ", nsites),
         `Low skew` = paste0(signif(100 * low_skew, 2), "%; n = ", nsites)) %>%
  select(-nsites)

metric_results <- sub_di %>%
  filter(nparts > 40) %>%
  group_by(Dataset) %>%
  summarize(high_simpson = mean(simpson_percentile_excl > 97.5),
            low_simpson = mean(simpson_percentile < 2.5),
            high_shannon = mean(shannon_percentile_excl > 97.5),
            low_shannon = mean(shannon_percentile < 2.5),
            high_nsingletons = mean(nsingletons_percentile_excl > 97.5),
            low_nsingletons = mean(nsingletons_percentile < 2.5),
            nsites = dplyr::n()) %>%
  ungroup() %>%
  mutate(`High Simpson` = paste0(signif(100 * high_simpson, 2), "%; n = ", nsites),
         `Low Simpson` = paste0(signif(100 * low_simpson, 2), "%; n = ", nsites),
         `High Shannon` = paste0(signif(100 * high_shannon, 2), "%; n = ", nsites),
         `Low Shannon` = paste0(signif(100 * low_shannon, 2), "%; n = ", nsites),
         `High proportion of rare species` = paste0(signif(100 * high_nsingletons, 2), "%; n = ", nsites),
         `Low proportion of rare species` = paste0(signif(100 * low_nsingletons, 2), "%; n = ", nsites)) %>%
  select(-nsites)

all_results <- left_join(diss_results, skew_results) %>%
  left_join(metric_results)

cols1 <- c("Dataset", "High dissimilarity", "High proportion of rare species", "High skew", "Low Simpson", "Low Shannon")

all_results %>%
  select(cols1)

# Very small communities

all_di <- all_di %>%
  mutate( `Number of elements` = ifelse(nparts < 2000, "Less than 2000", "More than 2000"))
diss_results <- all_di %>%
  filter(nparts > 20) %>%
  group_by(Dataset,  `Number of elements` ) %>%
  summarize(high_diss = mean(real_po_percentile_excl > 95),
            nsites = dplyr::n()) %>%
  ungroup() %>%
  mutate(`High dissimilarity` = paste0(signif(100 * high_diss, 2), "%; n = ", nsites)) %>%
  select(-nsites)

skew_results <- all_di %>%
  filter(nparts > 40, s0 > 2) %>%
  group_by(Dataset,  `Number of elements` ) %>%
  summarize(high_skew = mean(skew_percentile_excl > 97.5),
            low_skew = mean(skew_percentile < 2.5),
            nsites = dplyr::n()) %>%
  ungroup() %>%
  mutate(`High skew` = paste0(signif(100 * high_skew, 2), "%; n = ", nsites),
         `Low skew` = paste0(signif(100 * low_skew, 2), "%; n = ", nsites)) %>%
  select(-nsites)

metric_results <- all_di %>%
  filter(nparts > 40) %>%
  group_by(Dataset,  `Number of elements` ) %>%
  summarize(high_simpson = mean(simpson_percentile_excl > 97.5),
            low_simpson = mean(simpson_percentile < 2.5),
            high_shannon = mean(shannon_percentile_excl > 97.5),
            low_shannon = mean(shannon_percentile < 2.5),
            high_nsingletons = mean(nsingletons_percentile_excl > 97.5),
            low_nsingletons = mean(nsingletons_percentile < 2.5),
            nsites = dplyr::n()) %>%
  ungroup() %>%
  mutate(`High Simpson` = paste0(signif(100 * high_simpson, 2), "%; n = ", nsites),
         `Low Simpson` = paste0(signif(100 * low_simpson, 2), "%; n = ", nsites),
         `High Shannon` = paste0(signif(100 * high_shannon, 2), "%; n = ", nsites),
         `Low Shannon` = paste0(signif(100 * low_shannon, 2), "%; n = ", nsites),
         `High proportion of rare species` = paste0(signif(100 * high_nsingletons, 2), "%; n = ", nsites),
         `Low proportion of rare species` = paste0(signif(100 * low_nsingletons, 2), "%; n = ", nsites)) %>%
  select(-nsites)

all_results <- left_join(diss_results, skew_results) %>%
  left_join(metric_results)

cols1 <- c("Dataset", "Number of elements", "High dissimilarity", "High proportion of rare species", "High skew", "Low Simpson", "Low Shannon")

all_results %>%
  select(cols1)

plot_narrowness_hist(filter(all_di, dat %in% c("fia", "mcdb", "misc_abund")), "sim_pos_from_best", "Dissimilarity") +
  facet_wrap(vars(`Number of elements`, Dataset), scales = "free_y", nrow = 2, ncol = 3)

plot_percentile_hist(filter(all_di, dat %in% c("fia", "mcdb", "misc_abund")), "real_po_percentile_mean", "Dissimilarity") +
  facet_wrap(vars(`Number of elements`, Dataset), scales = "free_y", nrow = 2, ncol = 3)

plot_narrowness_hist(filter(all_di, dat %in% c("fia", "mcdb", "misc_abund")), "nsingletons_95_ratio_2t", "Number of rare species") +
  facet_wrap(vars(`Number of elements`, Dataset), scales = "free_y", nrow = 2, ncol = 3)

plot_percentile_hist(filter(all_di, dat %in% c("fia", "mcdb", "misc_abund")), "nsingletons_percentile_mean", "Number of rare species") +
  facet_wrap(vars(`Number of elements`, Dataset), scales = "free_y", nrow = 2, ncol = 3)

plot_narrowness_hist(filter(all_di, dat %in% c("fia", "mcdb", "misc_abund")), "simpson_95_ratio_2t", "Simpson") +
  facet_wrap(vars(`Number of elements`, Dataset), scales = "free_y", nrow = 2, ncol = 3)

plot_percentile_hist(filter(all_di, dat %in% c("fia", "mcdb", "misc_abund")), "simpson_percentile_mean", "Simpson") +
  facet_wrap(vars(`Number of elements`, Dataset), scales = "free_y", nrow = 2, ncol = 3)

plot_narrowness_hist(filter(all_di, dat %in% c("fia", "mcdb", "misc_abund")), "shannon_95_ratio_2t", "Shannon") +
  facet_wrap(vars(`Number of elements`, Dataset), scales = "free_y", nrow = 2, ncol = 3)

plot_percentile_hist(filter(all_di, dat %in% c("fia", "mcdb", "misc_abund")), "shannon_percentile_mean", "Shannon") +
  facet_wrap(vars(`Number of elements`, Dataset), scales = "free_y", nrow = 2, ncol = 3)

ggplot(all_di, aes(s0, n0, color = `Number of elements`)) +
  geom_point() +
  xlim(0, 30) +
  ylim(0,150)

all_di <- all_di %>%
  mutate( `Number of elements` = ifelse(nparts < 1000, "Less than 1000", "More than 1000"))
diss_results <- all_di %>%
  filter(nparts > 20) %>%
  group_by(Dataset,  `Number of elements`) %>%
  summarize(high_diss = mean(real_po_percentile_excl > 95),
            nsites = dplyr::n()) %>%
  ungroup() %>%
  mutate(`High dissimilarity` = paste0(signif(100 * high_diss, 2), "%; n = ", nsites)) %>%
  select(-nsites)
ggplot(filter(all_di, dat == "mcdb"), aes(skew_95_ratio_2t)) + 
  geom_histogram() +
  facet_wrap(vars(`Number of elements`), scales = "free_y")

skew_results <- all_di %>%
  filter(nparts > 40, s0 > 2) %>%
  group_by(Dataset,  `Number of elements`) %>%
  summarize(high_skew = mean(skew_percentile_excl > 97.5),
            low_skew = mean(skew_percentile < 2.5),
            nsites = dplyr::n()) %>%
  ungroup() %>%
  mutate(`High skew` = paste0(signif(100 * high_skew, 2), "%; n = ", nsites),
         `Low skew` = paste0(signif(100 * low_skew, 2), "%; n = ", nsites)) %>%
  select(-nsites)

metric_results <- all_di %>%
  filter(nparts > 40) %>%
  group_by(Dataset, `Number of elements`) %>%
  summarize(high_simpson = mean(simpson_percentile_excl > 97.5),
            low_simpson = mean(simpson_percentile < 2.5),
            high_shannon = mean(shannon_percentile_excl > 97.5),
            low_shannon = mean(shannon_percentile < 2.5),
            high_nsingletons = mean(nsingletons_percentile_excl > 97.5),
            low_nsingletons = mean(nsingletons_percentile < 2.5),
            nsites = dplyr::n()) %>%
  ungroup() %>%
  mutate(`High Simpson` = paste0(signif(100 * high_simpson, 2), "%; n = ", nsites),
         `Low Simpson` = paste0(signif(100 * low_simpson, 2), "%; n = ", nsites),
         `High Shannon` = paste0(signif(100 * high_shannon, 2), "%; n = ", nsites),
         `Low Shannon` = paste0(signif(100 * low_shannon, 2), "%; n = ", nsites),
         `High proportion of rare species` = paste0(signif(100 * high_nsingletons, 2), "%; n = ", nsites),
         `Low proportion of rare species` = paste0(signif(100 * low_nsingletons, 2), "%; n = ", nsites)) %>%
  select(-nsites)

all_results <- left_join(diss_results, skew_results) %>%
  left_join(metric_results)

cols1 <- c("Dataset", "Number of elements", "High dissimilarity", "High proportion of rare species", "High skew", "Low Simpson", "Low Shannon")

all_results %>%
  select(cols1)

# Resampling

all_di_with_resamples <- read.csv(here::here("analysis", "rev_prototyping", "all_di.csv"))

jk_di_full <- read.csv(here::here("analysis", "rev_prototyping", "all_di_jk.csv"))

all_ct_with_resamples <- read.csv(here::here("analysis", "rev_prototyping", "all_ct.csv"))

jk_ct <- read.csv(here::here("analysis", "rev_prototyping", "all_ct_jk.csv"))

jk_di_full <- left_join(jk_di_full, jk_ct)

all_di_with_resamples <- left_join(all_di_with_resamples, all_ct_with_resamples)

jk_di <- jk_di_full %>%
  group_by_all() %>%
  mutate(site_source = unlist(strsplit(site, "_")[[1]][[1]])) %>%
  ungroup() %>%
  select(dat, site, site_source, s0, n0, skew_percentile_excl,skew_percentile, simpson_percentile, simpson_percentile_excl, shannon_percentile, shannon_percentile_excl, nsingletons_percentile, nsingletons_percentile_excl, real_po_percentile, real_po_percentile_excl, sim_pos_from_best, skew_95_ratio_2t, simpson_95_ratio_2t, shannon_95_ratio_2t, nparts)

skewcols <- colnames(jk_di)[ grepl("skew", colnames(jk_di))]

jk_di[ which(jk_di$s0 < 3), skewcols] <- NA

jk_di_means <- jk_di %>%
  filter(n0 != s0,
         s0 != 1,
         n0 != (s0 + 1)) %>%
  group_by(dat, site_source) %>%
  mutate(njks = dplyr::n(),
         njks_skew = sum(!is.na(skew_percentile)))%>%
  summarise_at(vars(-group_cols(), -site), mean, na.rm =  T) %>%
  mutate(resampling = "Subsampling") %>%
  mutate(site = as.numeric(site_source)) %>%
#  select(-site_source) %>%
  mutate(dat = ifelse(grepl(dat, pattern = "fia"), "fia", dat),
         dat = ifelse(dat == "misc_abund_short", "misc_abund", dat)) %>%
  mutate(Dataset = dat,
         Dataset = ifelse(Dataset == "fia", "FIA", Dataset),
         Dataset = ifelse(Dataset == "bbs", "Breeding Bird Survey", Dataset),
         Dataset = ifelse(Dataset == "mcdb", "Mammal Communities", Dataset),
         Dataset = ifelse(Dataset == "gentry", "Gentry", Dataset),
         Dataset = ifelse(Dataset == "misc_abund", "Misc. Abundance", Dataset)) %>%
  filter(nparts > 20,
         njks == 10) %>%
  ungroup()

jk_di[ which(jk_di$njks_skew < 10), skewcols] <- NA

skewcols <- colnames(all_di_with_resamples[ which(grepl("skew", colnames(all_di_with_resamples)))])

all_di_with_resamples[ which(all_di_with_resamples$s0 < 3), skewcols] <- NA

all_di_with_resamples <- all_di_with_resamples %>%
  mutate(log_nparts = log(gmp:::as.double.bigz(nparts)),
         log_nsamples = log(nsamples),
         log_s0 = log(s0),
         log_n0 = log(n0)) %>%
  filter(n0 != s0,
         s0 != 1,
       #  !singletons,
         n0 != (s0 + 1)) %>%
  mutate(dat = ifelse(grepl(dat, pattern = "fia"), "fia", dat),
         dat = ifelse(dat == "misc_abund_short", "misc_abund", dat)) %>%
  mutate(Dataset = dat,
         Dataset = ifelse(Dataset == "fia", "FIA", Dataset),
         Dataset = ifelse(Dataset == "bbs", "Breeding Bird Survey", Dataset),
         Dataset = ifelse(Dataset == "mcdb", "Mammal Communities", Dataset),
         Dataset = ifelse(Dataset == "gentry", "Gentry", Dataset),
         Dataset = ifelse(Dataset == "misc_abund", "Misc. Abundance", Dataset)) %>%
  filter(nparts > 20) %>%
  mutate(resampling = ifelse(singletons, "Rare species", "Raw")) %>%
  bind_rows(jk_di_means)

>>>>

---
title: "Distribution of dissimilarity and breadth indices within each dataset"
author: Renata Diaz
date: "`r Sys.Date()`"
output: 
    # github_document:
    #    df_print: kable
    #    toc: true
    word_document:
        df_print: kable
        toc: false
---

@@

@@

  ggplot(di_df, aes(nparts, response, color = Dataset)) +
    geom_point(data = filter(di_df, dat == "fia"), alpha = .1) +
    geom_point(data = filter(di_df, dat != "fia"), alpha = .1) +
    geom_point(data = filter(di_df, s0 < 0)) +
    scale_color_viridis_d(end = .9) +
    xlab("") +
    ylab(ylabel) +
    scale_x_log10() +
    ggtitle(yvar_name) +
    theme(legend.position = "none")+
    theme(plot.title = element_text(size=10))
}

@@

@@

**Figure S6**. Partially because of the uneven distribution of S and N among the different datasets, the narrowness of the feasible sets - defined either as the mean dissimilarity of samples from the feasible set to the central tendency of the feasible set, or using a breadth index for specific metrics - varies among different datasets. In particular, the FIA dataset, and subsets of the Mammal Community and Miscellaneous Abundance databases, often have highly variable, broadly-defined statistical baselines derived from the feasible set.
